Comparing some optimization methods to obtain estimations in the Gompertz-lognormal diffusion process

### 2015-01-23 paper.tex
Updating title and abstract for GECCO

\begin{abstract}
The maximum likelihood estimation of the parameters of a diffusion process requires a system of equations that, for some cases, has no explicit solution to be solved. Facing that situation, we can approximate the solution using an optimization method. In this paper we compare five optimization methods: an Iterative Method, an algorithm based on Newton-Raphson solver, a Variable Neighbourhood Search method, a Simulated Annealing algorithm and an Evolutionary Algorithm. We generate four data sets following a Gompertz-lognormal diffusion process using different noise level. The methods are applied with these data sets for estimating the parameters which are present into the diffusion process. Results show that bio-inspired methods gain suitable solutions for the problem every time, even when the noise level increase. On the other hand, some analytical methods as Newton-Rapson or the Iterative Method do not solve always solve the problem whether their scores depend on the starting point for initial solution or the noise level hinders the resolution of the problem. In these cases, the bio-inspired algorithms remain as a suitable and reliable approach.

%Maribel: he modificado ligeramente la primera frase para generalizar un poco el problema y que no parezca desde el principio una cosa tan sumamente específica. En esta generalización, es posible que haya dicho algo que no es verdad, así que por favor, si es así, intentad corregirlo, pero sin especificar mucho, para que el paper no parezca desde el principio algo muy muy específico y poco generalizable, aunque lo sea.

>>>>

\begin{abstract}
The maximum likelihood estimation of the parameters of a diffusion
process requires a system of equations that, for some cases, has no
explicit solution; thus, the solution has to be estimated. We will use
optimization methods to find those solutions: an Iterative Method, an
algorithm based on Newton-Raphson solver, a Variable Neighbourhood
Search method, a Simulated Annealing algorithm and an Evolutionary
Algorithm. We generate four data sets following a Gompertz-lognormal
diffusion process using different noise levels. The methods are applied
to these data sets in order to estimate the parameters that rule
the diffusion process. Results show that bio-inspired methods
yield suitable solutions for the problem every time, even when the
noise level increases. On the other hand, some analytical methods as
Newton-Raphson or the Iterative Method do not always solve the
problem whether their scores depend on the starting point for initial
solution or the noise level hinders the resolution of the problem. In
these cases, the bio-inspired algorithms remain as a suitable and
reliable approach.
